---
title:  "Case Study: Expanding Our Candy Brand"
author:
- Nandhini Santhanam
output:
  html_document:
    df_print: paged
    toc: yes
    toc_depth: '3'
  pdf_document:
    number_sections: yes
    toc: yes
    toc_depth: 2
latex_engine: xelatex
always_allow_html: true
abstract: |
    The lidl purchasing group intends to introduce a new candy product to increase their sales and attract new customers.This work intends to identify the characteristics of different chocolates available in the market which make them special and popular among people and thereby provide a business recommendation for the new product. In order to achieve this goal, the market research data on Candy power ranking from fivethirtyeight database is utilized.
---



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



# Introduction
```{r imports, echo=TRUE, message=FALSE, warning=FALSE}
#Library needed for different purpose
library(plyr)
library(dplyr)
library(skimr)
library(plotly)
library(corrplot)
library(factoextra)
library(NbClust)
```


# Problem Understanding

The main aim of this work is to provide a business recommendation on the different characteristics of the candies that are more likely to become popular among the people. Using these insights, a new candy brand can be introduced to have a profit for the lidl business group.
This can be achieved by converting the problem statement to a data science objective This is then followed by different analysis. This flow is explained in this document and finally a business recommendation is provided.

# Data Understanding

This section consist of the details about the data set that is used for the analysis and also the details about  the quality of the data such as information about missing values.

### Overview About The Data

The input csv file consisted of 13 variables and 85 observations. The variables consists of the competitor name, 9 Boolean variable which indicates the various characteristics of the chocolate that is surveyed, percentile of sugar content in the chocolate when compared within the data set, percentile of unit price when compared with the rest of the data in the set and the overall win percentage from the total of 269,000 match ups. These can be associated with the column names that is shown below.


```{r}
#reading the input csv file
Candy_data <- read.csv("D:\\RstudioProjects\\candy-power-ranking\\candy-data.csv")
print(colnames(Candy_data))
dim(Candy_data)

```


### Quality Of The Data

When analyzing the type of the data, the winpercent,sugarpercent and pricepercent are found to be in character. It makes sense to have these quantities as numeric to do further calculation and analysis using it.The result is shown in the below code block.

```{r Data quality}
#checking data type of each column
apply(Candy_data,MARGIN = 2,FUN = class)

#converting the appropriate column to numeric
Candy_data$sugarpercent <- as.numeric(Candy_data$sugarpercent)
Candy_data$pricepercent <- as.numeric(Candy_data$pricepercent)
Candy_data$winpercent <- as.numeric(Candy_data$winpercent)

#Candy_data[2:13] <- as.numeric(Candy_data[2:13])
#check if the data type is updated 

print("After the updation")
apply(Candy_data[11:13],MARGIN = 2,FUN = class)

```
Skim function from package skimr allows you to comprehensively check the quality of the data such as information about missing values. The data that is used for the analysis of this work does not have any missing value which is shown by the n\_missing attribute for every column of the original data frame.


```{r Check for missing values}
#Check for missing values, statistical value like mean,median,sd for numerical quantities.
skimmed_Info <- skim(Candy_data)

#check if any column has missing values

which(skimmed_Info$n_missing > 0)
```


We can also skim through the numeric features such as __winpercent , pricepercent, sugarpercent__. This function shows the statistical values such as mean,median,quantiles and also its histogram in the table below. 
```{r analyse each of the numeric column winpercent}
# skimming of WinPercent
skim(Candy_data) %>%
  dplyr::filter(skim_variable == "winpercent")
```


```{r analyse each of the numeric column pricepercent}
# skimming of pricepercent
skim(Candy_data) %>%
  dplyr::filter(skim_variable == "pricepercent")
```




```{r analyse each of the numeric column sugarpercent}
# skimming of sugarpercent
skim(Candy_data) %>%  dplyr::filter(skim_variable == "sugarpercent")

```



# Objective - Data Science Problem 

In order to achieve the business recommendation, data driven analytic approach should be followed. Using the data set provided, beneficial insights must be derived. In order to achieve the same, the following tasks are carried out and a final conclusion in formulated based on its results.

1. Plotting of histogram of the different variables and arrive at a statistical   inference about its distribution.
2. Analyse the correlation between different variables in the data set with that of the win percentage.
3. Perform exploratory analysis to identify the different behavior and pattern in the data.
4. Perform a cluster analysis to identify if there are any clusters present in the data set.


# Exploratory Analysis

## Histogram Of The Variables

In the data set there are 9 quality variables which has values either 1 or 0. 1 indicates the corresponding characteristics to be present in the candy and 0 indicates the corresponding characteristics is not present in the candy.
In order to find the distribution of these variables, a histogram is plotted for these variables using Plotly package. The code below shows the same.


```{r histogram}
Candy_data[,c(2:10)]<-apply(Candy_data[2:10],MARGIN = 2,FUN = as.factor)


plotList = list()
plotListNames = colnames(Candy_data[2:10])
```


```{r histogram function stacked}

#function to plot the histograms of different variables
Stackedhistogram <- function(plotListNames ,Candy_data)
{
  plotList = list()
for(i in 1:length(plotListNames))
{
thisName = plotListNames[i]
plotList[[thisName]] <-plotly_build( plot_ly(x=~Candy_data[,i],type="histogram", name = thisName) %>%  layout(title="Histogram of different variables",xaxis = list(
title = plotListNames[i]),yaxis = list(title = "Count")) )
  
}
return(plotList)  
}

plotList<-Stackedhistogram(plotListNames,Candy_data[2:10])


subplot(plotList,nrows = 3,shareY = T)

```


* It can be seen that the characteristics like __caramel__, __peanutalmondy__, __nougat__, __crispedricewafer__, __hard__ and __bar__ are mostly not present in many of the chocolates. This is indicated by the high count of these variables having a value of 0. 
* The candies with nougat, crispedricewafer are less popular candies
* The characteristics like __chocolate__, __fruity__ and __pluribus__ are having probability of getting 1 and 0 almost close to each other. But the count of getting 0 in chocolate and caramel is little high than having a value of 1. whereas in pluribus the count of having a 1 is little more than that of probability of having a 0.
* This shows that the many candies are coming as pack of many individual candies(Pluribus). There is a equal balance of fruit and chocolate candies in the data set.

## Correlation Analysis

Correlation analysis will provide more insights to determine what type of characteristics has good correlation with that of the other features. A correlation matrix is formed between all the variables leaving out the competitor name.

```{r correlation matrix}
Candy_data[,c(2:10)]<-apply(Candy_data[2:10],MARGIN = 2,FUN = as.numeric)
M <- cor(x = Candy_data[2:13])
```

In order to determine the statistical significance of the correlation co efficient, p value is calculated for the correlation matrix. This is calculated by the function which is mentioned in the below code block.

```{r function to calculate p value of correlation matrix}
#function to calculate p value of correlation matrix

cor.mtest <- function(mat, ...) {
    mat <- as.matrix(mat)
    n <- ncol(mat)
    p.mat<- matrix(NA, n, n)
    diag(p.mat) <- 0
    for (i in 1:(n - 1)) {
        for (j in (i + 1):n) {
            tmp <- cor.test(mat[, i], mat[, j], ...)
            p.mat[i, j] <- p.mat[j, i] <- tmp$p.value
        }
    }
  colnames(p.mat) <- rownames(p.mat) <- colnames(mat)
  p.mat
}


# matrix of the p-value of the correlation
p.mat <- cor.mtest(Candy_data[2:13])

```

The below plot shows the correlation between different variables with the associated value in each of the square. The correlation coefficient value whose p value is less than 0.05 is marked with a X mark.

```{r correlation plot}
corrplot(M, method="color",  
         type="upper", order="hclust", 
         addCoef.col = "black", # Add coefficient of correlation
         tl.col="black", tl.srt=45, #Text label color and rotation
         # Combine with significance
         p.mat = p.mat, sig.level = 0.05, 
         # hide correlation coefficient on the principal diagonal
         diag=F,title = "Correlation Plot between the variables",
           mar=c(0,0,1,0) 
         )
```

* Chocolate and fruity characteristics have high negative correlation indicating that there is a very less chance that the candy has both chocolate and fruit in it.
* There is only one such data point in the data set and it its __Tootsie Pop__.

```{r candy that has both chocolate and fruity flavour in it, echo=FALSE}
Candy_data[which(Candy_data$chocolate ==1 & Candy_data$fruity==1),c(1:3,11:13)]

```
We can derive the correlation of the other input variables with respect to the winpercentage variable alone.

```{r}
print(cor(Candy_data[2:12],Candy_data$winpercent))
```
* The above result indicates that the __chocolate__ has high positive correlation with that of the win percent indicating the high popularity with customers.

* Even bar, peanut almondy,  crispedricewafer has positive correlation with the win percent. But the number of observations that have a value of one for these characteristics are less in the total observations. This was also observed in the histogram plot as well.

* The __fruity__ and __pluribus__ has negative correlation with that of win percent indicating that these are not popular among customers.

* The hard type also has a negative correlation with the win percent but the number of observation with one value in hard type is less when compared to total number of observation.


## Derived Features
1. __Type__ : A new feature is calculated to find out the unique combinations of different characteristics that are present in the candies. In order to find it, the binary combinations of 1 and 0 is converted to a decimal value. This method gives a unique value for every combination.
```{r Derived features - 1}
#Type id to indicate the unique combinations [Conversion of binary to decimal approach]
Candy_data$Type <- (Candy_data$chocolate* 1) + (Candy_data$fruity* 2) +  (Candy_data$caramel * 4) + (Candy_data$peanutyalmondy* 8) + (Candy_data$nougat* 16) + (Candy_data$crispedricewafer* 32) + (Candy_data$hard* 64) + (Candy_data$bar* 128) + (Candy_data$pluribus* 256)

summary(Candy_data$Type)

#length of unique combinations
length(unique(Candy_data$Type))

#frequency of each type of combinations

Unique_Combinations <-plyr::count(Candy_data$Type)

print(Unique_Combinations[c(order(Unique_Combinations$freq,decreasing = T)),])
```
+ The above feature shows that there are 29 unique combinations from that of the 95 observations.
+ The feature type 258 (Pluribus - 1 and Fruity - 1) is present the most number of times i.e. 19 times.
+ The second most repeated type is 322 (Pluribus - 1, Hard - 1 and Fruity - 1) and it had appeared 7 times.
+ The first three most repeated type has Pluribus as 1.


2. __NoOfQualityFeatures__ : This feature indicates the total number of characteristics which is present in a single competitor. This is calculated by adding the characteristics features.
```{r Derived features - 2}
#total number of qualitative characteristics available for a competitor.
Candy_data$NoOfQualityFeatures <- apply(Candy_data[2:10],MARGIN = 1,FUN = sum)

#summary

summary(Candy_data$NoOfQualityFeatures)
```

```{r Histgram for No of features}

#histogram for NoOfQualityFeatures
plot_ly(x=~Candy_data[,15],type="histogram") %>%  layout(title="Histogram of the Number of quality features",xaxis = list(
title = "NoOfQualityFeatures"),yaxis = list(title = "Count")) 
```

+ The histogram plot shows that the most of the candies have either 2 or 3 characteristics.

+ There are two candies which does not have any of the characteristics that is listed in the data set and they are "one dime" and "one quarter". These two have the same sugar percent also. So the winpercent for these two candies solely depends on the pricepercent value. obviously the type variable is also 0 for both the candies.This is indicated by the correlation between them to be 1.

```{r code for the same}

cor(Candy_data$pricepercent[which(Candy_data$NoOfQualityFeatures==0)],Candy_data$winpercent[which(Candy_data$NoOfQualityFeatures==0)])
```
## Observations From Data

1. Top ten winner candies based on its win percentile

```{r list of 10 winner candies based on its winpercentile}
Top_10_Candy<-head(Candy_data[c(order(Candy_data$winpercent,decreasing = T)),],10)
Top_10_Candy
```
+ All the top 10 candies are having the feature __chocolate__ as 1 and __fruity__ as 0. The candies that are chocolaty and not fruity are having high winpercentage among the customers.
+ The __candy type 9__ (Chocolate - 1 and PeanutAlmondy - 1)are present 3 times in the data set and all three competitors are present in the __top 10 candies__. Upon analyzing the correlation of sugarpercent and pricepercent of this type of candy with that of win percent shows that they are having negative correlation of -0.5 and -0.3. With only three data points its tough to conclude which affects the winpercent. But the type 9 which is the candies from Ressea is popular.
```{r}
#analysis on candy type 9
Candy_type_9 <- Candy_data[which(Candy_data$Type==9),c(1,11:14)]
head(Candy_type_9)
cor(Candy_type_9[2:3],Candy_type_9$winpercent)
```

+ In total there are only 14 candies that have the __peanutalmondy__ feature as 1 in the total of 85 observations. But in that, __7__ of them are in the top 10 candy list. This indicates that there is 50% chance that a candy having peanutalmondy to be more popular with the customer.
```{r}
#total number of candies having peanut almondy in the top 10 candies
length(which(Top_10_Candy$peanutyalmondy==1))

#total number of candies having peanut almondy in the whole data set
length(which(Candy_data$peanutyalmondy==1))
```
+ All the top 10 candies have __more than 1__ characteristics which can be seen from the "NoOfQualityFeatures" variable shown below.
```{r}
length(which(Top_10_Candy$NoOfQualityFeatures > 1))
```


+ When analyzing the pricepercent feature of the top 10 candies, 6 candies were having a price percentile value __0.651__ which is composed of different characteristics (different Type value).
+ There is a candy __Reeseas Miniatures__ in the top ten candies which is very cheap.
```{r echo=FALSE}
Top_10_Candy[2,c(1,11:14)]
```

+  There is a candy __Twix__ in the top ten candies that is very costly and also popular.

```{r echo=FALSE}
Top_10_Candy[3,c(1,11:14)]
```
+ These are some critical observations from the data.
# Cluster Analysis

Before starting with if there is a clusters present in the data, When there is a multicollinearity in the variables, it is better to do a dimensionality reduction and then perform the cluster analysis. The below code block shows the reduction of the input variable into two principal components.

```{r principal component reduction}
#reduction into two principal components
pca_data<-prcomp(Candy_data[,c(2:13)],rank. = 2,center = T,scale. = T)
# retriving the principal component values as a data frame
pca_data <- as.data.frame(pca_data$x)
```


This principal component is given as input to the cluster algorithm to determine the clusters present in the data. Kmean clustering is one type of clustering algorithm. This requires the number of clusters(K) to be formed in the data. 
Apt K value should be found to avoid overfitting or under fitting of the data.This is found using a package NbClust. This specifies the apt number of cluster to be formed in the data.


```{r Cluster Analysis, include=FALSE}
# NbClust package is used to determine the number of clusters
Clus_data <- NbClust(pca_data, min.nc=2, max.nc=5, method="kmeans")
```



The NbClust package suggest the Number of cluster to be __3__. This is then given as input to fit the Kmean algorithm with the two principal components as input data.
The clusters with ID 1,2,3 id formulated by the algorithm. This is then merged with the original data to proceed with further analysis.
```{r kmeans clustering}
#fitting Kmeans Cluster
fitKmeans<-kmeans(pca_data,centers = 3)

#concatenating the cluster ID with the PCA_data frame
pca_data$clusters <- fitKmeans$cluster

Candy_data <- cbind(Candy_data,pca_data)
```

```{r}
#plotting the principal components and coloured by the cluster numbers
plot_ly(data = Candy_data, x = ~PC2, y = ~PC1, color = ~as.factor(pca_data$clusters),type = "scatter",mode="markers",colors="Set1")%>%  layout(title="Scatter plot between PC's and coloured by Cluster ID",xaxis = list(
title = "Principal Component 2"),yaxis = list(title = "Principal Component 1"))
```

In order to analyse the characteristics of the three clusters, the below data frame is created.

```{r Analysis of the Clusters}
cluster_Data <- Candy_data %>% 
                group_by(clusters)%>%
                summarise(count=length(clusters),Chocolate=length(which(chocolate==1)),fruity=length(which(fruity==1)),caramel=length(which(caramel==1)),peanutyalmondy=length(which(peanutyalmondy==1)),nougat=length(which(nougat==1)),crispedricewafer=length(which(crispedricewafer==1)),hard=length(which(hard==1)),bar=length(which(bar==1)),pluribus=length(which(pluribus==1)),AvgSugar =mean(sugarpercent),AvgPrice = mean(pricepercent),Avgwin =mean(winpercent))

cluster_Data

```
* __Observations From The Cluster Analysis__
  + Almost all the Candies in the __clusters 1 and 2__ contains __chocolate__ in its ingredient and does not contain fruit.
  + Only Candies in __Cluster 3__ has __fruit__ in its ingredient.
  + Only Candies in __Cluster 2__ has __Nougat__ in its ingredient.
  + Only Candies in __Cluster 2__ is __hard__ in nature.
  + All the Candies in __Cluster 2__ is in the form of __bar__
  + The main difference between candies in cluster 1 and 2 are. Candies in __Cluster 1__ are in __Pluribus__ form and Candies in __Cluster 2__ are in the form of __bar__
  + Most candies in __Cluster 3__ are in __Pluribus__ form.
  + The candies (10 candies) that have both __Chocolate and Caramel__ fall in either of __Cluster 1 and 2__ (Top Clusters). The candies that have only caramel(4 Candies) will fall in __Cluster 3__.
```{r Chocolate and caramel}

#total count of caramel
length(which(Candy_data$caramel==1))
#total count of chocolate and caramel
length(which(Candy_data$caramel==1 & Candy_data$chocolate ==1))
#clusters of Candies with chocolate and caramel
(Candy_data$clusters[which(Candy_data$caramel==1 & Candy_data$chocolate ==1)])
#clusters of Candies without chocolate and with caramel
(Candy_data$clusters[which(Candy_data$caramel==1 & Candy_data$chocolate !=1)])
```


+ There is no much difference between these clusters when its comes to Average Sugar percent in it.
+ Candies in __Cluster 2__ are the __Costliest__ and the candies in __Cluster 3__ are the __Cheapest__.
+ Cluster 2 has the __highest win percentage(62.6%)__ and the __Cluster 1__ has the __second highest win percentage(60.4%)__ and __Cluster 3__ is the __least__.



# Business Recommendation From The Analysis
Based on the exploratory and cluster analysis, the following business recommendation can be useful in achieving the popularity for newly introducing candy brand.

* Candies that are chocolaty are more popular with the customers.
* Based on data, if the caramel comes with the chocolate flavor rather than alone is famous with customer
* Candies having Peanutalmondy is having 50% chance to end up in top 10 candies in the market.
* Candies that fall in the Cluster 1 and Cluster 2 are having almost similar  average win percentage of around 61%. 
* But the Candies that fall in Cluster 2 has high average price percent when compared to candies that fall in Cluster 1. Even though the candies in Cluster 2 is costlier, they are having high win percentage.This (Candies of Cluster 2) can be little more economically beneficial to the lidl business group when compared to the Candies of Cluster 1 if their manufacturing cost is similar.
* Introducing candies with characteristics that of the candies in Cluster 1 and 2 would be the best for Lidl business group.
  + Candies with a layer of __Chocolate__ and __Caramel__ 
  + Candies with a __PeanutAlmondy__ in it.
  + Candies with __Nougat__ in it.
  + Candies should __not be hard__ 
  + Candies should come in __Bar__ form.

# Conclusion
1. Thus the objective of the work to provide a business recommendation about the characteristics of the candy that should be introduced by the lidl group to become popular among customers are achieved.
2. The solution is achieved by first setting the objective in data science perspective and further dividing into sub tasks.
3. The results from the various sub tasks were useful in arriving at the business recommendation.

